Drought regimens predict life history strategies in \emph{Heliophila}

### 2019-06-08 manuscript/manuscript.tex
proofing

\shorttitle{Drought and life history}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{2} College of Agriculture, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{3} Institute for Environment and Society, Brown University, Providence, RI 02912, USA\\\textsuperscript{4} Biology Department, Pennsylvania State University, State College, PA 16802, USA}
\keywords{drought adaptation, herbaria records, Heliophila, life history evolution, phylogeography, remote sensing }
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

Understanding the causes and consequences of life history variation is a
longstanding goal of ecology and evolutionary biology (Cole, 1954). In
plants, life histories are especially diverse, with some species
completing their life cycle in a number of weeks to others that live for
thousands of years (Brown, 1996). Along this continuum in angiosperms an
important division exists distinguishing annuals which complete their
seed to seed life cycle within a single calendar year from perennials
which can persist over multiple years. Annual plants do not need to
survive through the full range of seasonal environmental variation and
spend at least some portion of the year as a seed where they are
relatively protected from environmental stress. In contrast, perennial
plants can continue vegetative growth over multiple years and must
survive conditions experienced during all seasons but can also benefit
from competitive advantages and, if iteroparous, multiple bouts of
reproduction. These represent fundamentally different life history
strategies and predicting their occurrence is important for community,
ecosystem, and agricultural ecology. However, the environmental factors
that explain their evolution and distributions remain empirically
unresolved (Friedman \& Rubin, 2015).

Classical theory predicts shorter life spans in environments where adult
mortality is high (Charnov \& Schaffer, 1973; Stearns, 1992; Franco \&
Silvertown, 1996). Because lack of water is perhaps the greatest threat
to survival during vegetative or reproductive growth in plants, this
theory has been extended to the hypothesis that annuality is adaptive
when it allows plants to escape drought (Schaffer \& Gadgil, 1975).
Indeed, adaptation to drought, defined as episodes of increased aridity
causing plant sress (Passioura, 1996), is often invoked as an
explanation for the success of annual species. And while a few cases are
cited where annuality appears to be more common in environments with
greater aridity (Stebbins Jr, 1952; Morishima \emph{et al.}, 1984), this
hypothesis has yet to be supported while controlling for the effect of
common ancestry (phylogeny) on life habit. In one previous study where
this question was addressed phylogenetically, (Evans \emph{et al.},
2005) annuals were not found to be associated with environments that
experience more drought. This could be explained by the relatively small
number of species studied and the reliance on a limited number of
weather stations to characterize environments, highlighting the need to
develop more scalable methods to study the geographic distributions of
traits such as life history. Thus, in this study we leverage thousands
of herbaria specimens among dozens of species and high-resolution remote
sensing to study the distributions and environmental factors potentially
driving the evolution and distribution of life history.

It is also critical to consider another dimension of drought adaptation:
the expectation that annuality is most adaptive when droughts are not
only frequent but also predictable. That is, when the frequency of
drought is particularly high during certain seasons. Such predictability
is important for selection to favor and escape strategy during those
seasons which are particularly drought prone. While there has been at
least one example of annuality associated with environments
qualitatively classified as \enquote{predictable} in a general sense
(Datson \emph{et al.}, 2008), the seasonal predictability of drought
experienced by annuals has yet to be rigorously studied. As such,
further empirical work is needed to support the model of annuality as a
mechanism of drought adaptation via escape from drought prone seasons.
Here we study herbarium collection dates to ask whether annuals indeed
exhibit evidence of an escape strategy from seasons with elevated
drought frequency.

@@

Here we combine a long-term global dataset of satellite detected drought
events with metadata from natural history collections to test these
classic hypotheses within the African endemic mustard genus,
\emph{Heliophila} L. (Brassicaceae). If annuality is an adaptive
strategy allowing plants to escape drought prone seasons, then drought
frequency should predict the distribution of life history strategies
across landscapes, and annual species should be more commonly associated
with drought prone regions than perennial species. Additionally, if
perenniality offers competitive advantage in the absence of drought,
associations between life history and drought frequency should be
significant when phylogenies include transitions from annual to
perennial life history strategy. Finally, if annual species have adapted
to escape predictably drought prone seasons, observations of growing
annual species (i.e.~occurring in forms other than seed) should be rare
during seasons when drought frequency is highest. Phylogenetic
relatedness can influence tests of associations between species' traits
and their environments by confounding common environments caused by
selection from common environments caused my ancestry. (Felsenstein,
1985; Barrett \emph{et al.}, 1996). Therefore, we assessed the
relationship between life history distribution and drought frequency
while controlling for phylogeny.

@@

To characterize drought regimens across the distributions of annual and
perennial species of \emph{Heliophila}, we calculated drought during
different seasons at the location of observations for \emph{Heliophila}
records using the VHI. Specifically, we created maps of the frequencies
of observing drought conditions between years (VHI\textless{}40, NOAA)
during the winter (quarter surrounding winter solstice), spring (quarter
surrounding spring equinox), summer (quarter surrounding summer
solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015
across the range of \emph{Heliophila}. From these maps, the drought
frequency (the number of times drought is observed divided by the total
number of years, 34) during the winter, spring, summer, and fall were
extracted for the locations of all GBIF records.

@@

@@

To evaluate the hypothesis that annual and perennial life history
strategies reflect adaptations to alternative drought regimens, we
tested the corresponding prediction that the observed distributions of
annual and perennial \emph{Heliophila} species would be significantly
associated with historic drought frequency. We tested for a relationship
between drought frequency and life history, season, and their
interaction by analysis of variance while including species as a random
effect using the lme4 package in R ({\textbf{???}}) and compared annuals
and perennials using by Tukey adjusted post-hoc contrasts. We next
calculated the mean drought frequency during the winter, spring, summer
and fall for each species. Because shared evolutionary history of
closely related species can lead to spurious associations between traits
and environments (Felsenstein, 1985), we tested for a relationship
between life history strategy and drought frequency while controlling
for phylogeny using phylogenetic logistic regression (Ives \& Garland,
2010). This statistical approach is designed to control for the
confounding effects of common ancestry's influence on demographic
features such as geospatial relationships when addressing hypotheses
about the role of natural selection on trait distributions.

@@

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/phylogeny} \caption{Species and examples of herbaria specimens of
\emph{Heliophila} (a) Phylogeny and life history strategies of species
studied. Orange circles at branch tips mark annual species and blue
circles mark perennial species. At internal nodes, pie charts indicate
the estimated posterior probability of being annual versus perennial.
Example herbaria specimens accessed via GBIF of (a) \emph{H. minima},
(b) \emph{H. deserticola}, (c) \emph{H. coronopifolia} and (d) \emph{H.
ephemera}. Images (b,d,e) courtesy of The Bavarian Natural History
Collections (CC BY-SA 4.0) and (c) The London Natural History Museum (CC
BY 4.0). Links to images are found in the supplement.}\label{fig:phylogeny}
\end{figure}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/maps_boxplots} \caption{Locations of occurrence records of (a) annual and (b)
perennial \emph{Heliophila.} Drought frequency during the (c) winter,
(d) spring, (e) summer and (f) fall detected using the VHI. (g) Drought
frequencies during each season at the observation locations of annual
and perennial \emph{Heliophila} (Post-hoc contrasts annuals and
perennials from ANOVA, ** = p \textless{} 0.01).}\label{fig:mapsboxplots}
\end{figure}

@@

Out of 8670 \emph{Heliophila} GBIF records, 6634 were for species with
reported life history (Mummenhoff \emph{et al.}, 2005), 2856 had
geospatial data, 2833 did not have geospatial issues, 2684 were located
on pixels classified as land having drought measurements, 2543 were
located in Africa, 2192 were not duplicated. Thus, after all filtering
steps, 2192 records for 42 species (Figure \ref{fig:phylogeny}, Table
\ref{tab:speciesmeanstable}) passed for further analyses. The number of
samples varied between species, with a mean of 52.19 samples per
species. \emph{H. rigidiuscula} had the most records, 201, and \emph{H.
cornellsbergia} the fewest, 2 (Table \ref{tab:speciesmeanstable}).

There were clear visual differences between the distributions of the 960
annual and the 1232 perennial \emph{Heliophila} observation records (see
Figure \ref{fig:speciesmaps} for maps of individual species). While
annual species were generally found in the western regions of South
Africa and Namibia, primarily in the Cape Floristic Region and Succulent
Karoo (Figure \ref{fig:mapsboxplots}a), the occurrence of perennials
extended to the southern and eastern coast of South Africa (Figure
\ref{fig:mapsboxplots}b).

@@

We found that the frequency of drought was significantly higher at the
locations of occurrence records for annual species. When comparing
across all occurrence records (all records rather than species means,
Figure \ref{fig:mapsboxplots}g), a mixed-model analysis of variance
which included species as random effect revealed a significant
relationship between drought frequency and life history (p \textless{}
0.01), season (p \textless{} 0.01) and their interaction (p \textless{}
0.01) (Table \ref{tab:anovatable}). Post-hoc contrasts showed that the
frequency of drought was significantly higher at the location of annuals
during the summer (z ratio = 3.93, p = \textless{} 0.01), and fall (z
ratio = 4.06, p \textless{} 0.01). Because a comparison across all
occurrence records does not account for variation in the number of
records per species (Table \ref{tab:speciesmeanstable}) or species
relatedness (Figure \ref{fig:phylogeny}a), we also tested whether mean
drought frequency values of each species were significantly different
between annuals and perennials using phylogenetic logistic regression.
We found that the mean drought frequencies were significantly higher
(\(\alpha = 0.05\)) in annual species during the spring, summer, and
fall (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a,b). These
findings indicate that common ancestry alone does not explain
differences in the drought frequencies experienced between the
environments of annual and perennial \emph{Heliophila}.

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:modelstable}Phylogenetic logistic regressions between life history, and the mean drought frequency observed at specimen sites of Heliophila species the winter, spring, summer, and fall.}
\begin{tabular}{lll}
\toprule
Predictor & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{P}\\
\midrule
Intercept & 0.7231 & 0.6636\\
Winter drought freq. & -1.5452 & 0.7274\\ \midrule
Intercept & 5.0107 & 0.0534\\
Spring drought freq. & -12.9014 & 0.0464\\ \midrule
Intercept & 7.7093 & 0.0054\\
Summer drought freq. & -19.9056 & 0.0042\\ \midrule
Intercept & 7.0162 & 0.0082\\
Fall drought freq. & -20.8174 & 0.0067\\ \midrule
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Annual species were scored as 0 and perennial species as 1.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/line_and_dates} \caption{(a) Comparison (mean +- SE) of drought frequency across
seasons calculated at the occurrence locations of GBIF records of annual
and perennial species of \emph{Heliophila}. (b) Results from
phylogenetic logistic regression, where the line shows the model fit and
* = p \textless{} 0.05, ** = p \textless{} 0.01. Annuals were scored as
0 and perennials as 1. (c) Collection dates of GBIF records of annual
and perennial species of \emph{Heliophila} in relation to the
photoperiodic calendar where day 1 is intermediate to the fall equinox
and winter solstice.}\label{fig:lineplots}
\end{figure}

The preceding results indicate that annual species are found in
environments where droughts are significantly more frequent, especially
in the summer and fall. Classic life history theory hypothesizes that
annuality reflects adaptation to such environments because it allows
species to escape predictable stressful conditions. If this is the case,
we would expect that annuals spend the drought prone seasons of summer
and fall as seeds. To test this hypothesis, we compared the dates of
occurrence records between annual and perennial \emph{Heliophila}
species. The distributions reveal a considerable difference in the
timing of observation of these two life histories. In comparison to
perennials, which appear to be collected throughout the year, annuals
are almost exclusively observed during the winter and spring (Figure
\ref{fig:lineplots}b). The differences between the distribution of
collection dates were significant by all tests (ks.test D = 0.25, p
\textless{} 0.01; bartlett.test K2 = 503.18, p \textless{} 0.01) This is
consistent with a model of life history in which annual species flower
in the spring, set seed, senesce, and die before the summer. Thus, these
annual species are likely to remain dormant during the summer and fall,
when drought is the strongest predictor of the distributions of annual
and perennial life histories (Figure \ref{fig:lineplots}a).

@@

To test the hypothesis that annual and perennial plants reflect
adaptation to alternative drought environments we examined the landscape
distribution of life history strategies in the large and diverse mustard
genus, \emph{Heliophila}. Using metadata of 2192 occurrence records and
a 34 year dataset of satellite-detected droughts, we tested the
prediction that annual species are more often observed in drought-prone
locations than perennial species, when controlling for phylogenetic
relatedness. We found that drought frequency is significantly different
between the distributions of annual and perennial species, with annuals
being found in environments with more frequent drought, and that this
signal is strongest during the seasons when annuals are likely escaping
via seed dormancy. These results remain significant while controlling
for the phylogenetic relationships of \emph{Heliophila} species,
yielding support for the role that natural selection has played in
driving contemporary distributions of these alternatives strategies in
relation to drought regimens.

@@

In conclusion, we find strong support for classic life history theory
that predicts that annuality is adaptive in environments with frequency
and predictable droughts. We report evidence consistent with a life
history model in annuals in which they escape drought prone seasons
during the seed phase of their life cycle. Finally, we find evidence
that the distributions of perennial lineages may indicate a competitive
advantage in areas where droughts are infrequent. More broadly, this
work highlights the irreplaceable value of natural history collections
and demonstrates the power of combining such information with large
scale remote sensing data to address outstanding classic hypotheses in
ecology and evolution.

@@

@@

@@

\leavevmode\hypertarget{ref-dittberner2018natural}{}%
\textbf{\textnormal{Dittberner H}, \textnormal{Korte A},
\textnormal{Mettler-Altmann T}, \textnormal{Weber A}, \textnormal{Monroe
G}, \textnormal{Meaux J de}}. \textbf{2018}. Natural variation in
stomata size contributes to the local adaptation of water-use efficiency
in arabidopsis thaliana. \emph{bioRxiv}: 253021.

@@

@@

>>>>

\shorttitle{Drought and life history}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{2} College of Agriculture, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{3} Institute for Environment and Society, Brown University, Providence, RI 02912, USA\\\textsuperscript{4} Biology Department, Pennsylvania State University, State College, PA 16802, USA}
\keywords{drought adaptation, herbaria records, Heliophila, life history evolution, phylogeography, remote sensing}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

Understanding the causes of life history variation and climate
adaptation are longstanding goals of ecology and evolutionary biology
(Raunkiaer, 1905; Turesson, 1925; Clausen \emph{et al.}, 1948; Cole,
1954). In plants, life histories are especially diverse, with some
species completing their life cycle in a number of weeks to others that
live for thousands of years (Brown, 1996). Along this continuum in
angiosperms an important division exists distinguishing annuals which
complete their seed to seed life cycle within a single year from
perennials which can persist over multiple years. Annual plants do not
need to survive through the full range of seasonal environmental
variation and spend at least some portion of the year as a seed where
they are relatively protected from environmental stress. In contrast,
perennial plants can continue vegetative growth over multiple years and
must survive conditions experienced during all seasons but can also
benefit from competitive advantages and, if iteroparous, multiple bouts
of reproduction. These represent fundamentally different life history
strategies and predicting their occurrence is important for community,
ecosystem, and agricultural ecology. However, the environmental factors
that explain their evolution and distributions remain empirically
unresolved (Friedman \& Rubin, 2015).

Classical theory predicts shorter life spans in environments where adult
mortality is high (Charnov \& Schaffer, 1973; Stearns, 1992; Franco \&
Silvertown, 1996). Because lack of water is one of the greatest threats
to survival during vegetative or reproductive growth in plants, this
theory has been extended to the hypothesis that annuality is adaptive
when it allows plants to escape drought (Cody \emph{et al.}, 1975;
Schaffer \& Gadgil, 1975). Indeed, adaptation to drought, defined as
episodes of increased aridity causing plant stress (Passioura, 1996), is
often invoked as an explanation for the success of annual species,
though drought adapted perennials are also well known (Raunkiaer, 1905).
And while a few cases are cited where annuality appears to be more
common in environments with greater aridity (Stebbins Jr, 1952;
Morishima \emph{et al.}, 1984), this hypothesis has yet to be supported
while controlling for the effect of common ancestry (phylogeny) on life
habit. In one previous study where this question was addressed
phylogenetically, (Evans \emph{et al.}, 2005) annuals were not found to
be associated with environments that experience more drought. This could
be explained by the relatively small number of species studied and the
reliance on a limited number of weather stations to characterize
environments, highlighting the need to develop more scalable methods to
study the geographic distributions of traits such as life history. Thus,
in this study we leverage thousands of herbaria specimens among dozens
of species and high-resolution remote sensing to study the distributions
and environmental factors potentially driving the evolution and
distribution of life history.

It is also critical to consider another dimension of drought adaptation:
the expectation that annuality is most adaptive when droughts are not
only frequent but also predictable. That is, when the frequency of
drought is particularly high during certain seasons. Such predictability
is important for selection to favor an escape strategy during those
seasons which are particularly drought prone. While there has been at
least one example of annuality associated with environments
qualitatively classified as \enquote{predictable} in a general sense
(Datson \emph{et al.}, 2008), the seasonal predictability of drought
experienced by annuals has yet to be rigorously studied. As such,
further empirical work is needed to support the model of annuality as a
mechanism of drought adaptation via escape from drought prone seasons.
Here we examine herbarium collection dates to ask whether annuals indeed
exhibit evidence of an escape strategy from seasons with elevated
drought frequency.

@@

Here we combine a long-term global dataset of satellite detected drought
events with metadata from natural history collections to test these
classic hypotheses within the African endemic mustard genus,
\emph{Heliophila} L. (Brassicaceae). If annuality is an adaptive
strategy allowing plants to escape drought prone seasons, then drought
frequency should predict the distribution of life history strategies
across landscapes, and annual species should be more commonly associated
with drought prone regions than perennial species. Additionally, if
perenniality offers competitive advantage in the absence of drought,
associations between life history and drought frequency should be
significant when phylogenies include transitions from annual to
perennial life history strategy. Finally, if annual species have adapted
to escape predictably drought prone seasons, observations of growing
annual species (i.e.~occurring in forms other than seed) should be rare
during seasons when drought frequency is highest. Phylogenetic
relatedness can influence tests of associations between species' traits
and their environments by confounding common environments caused by
selection from common environments caused by ancestry. (Felsenstein,
1985; Barrett \emph{et al.}, 1996). Therefore, we assessed the
relationship between life history distribution and drought frequency
while controlling for phylogeny.

@@

To characterize drought regimens across the distributions of annual and
perennial species of \emph{Heliophila}, we calculated drought during
different seasons at the location of observations for \emph{Heliophila}
records using the VHI. Specifically, we created maps of the frequencies
of observing drought conditions between years (VHI\textless{}40, NOAA)
during the winter (quarter surrounding winter solstice), spring (quarter
surrounding spring equinox), summer (quarter surrounding summer
solstice) and autumn (quarter surrounding autumn equinox) from 1981 to
2015 across the range of \emph{Heliophila}. From these maps, the drought
frequency (the number of times drought is observed divided by the total
number of years, 34) during the winter, spring, summer, and autumn were
extracted for the locations of all GBIF records.

@@

Out of 8670 \emph{Heliophila} GBIF records, 6634 were for species with
reported life history (Mummenhoff \emph{et al.}, 2005), 2856 had
geospatial data, 2833 did not have geospatial issues, 2684 were located
on pixels classified as land having drought measurements, 2543 were
located in Africa, 2192 were not duplicated.

@@

To evaluate the hypothesis that annual and perennial life history
strategies reflect adaptations to alternative drought regimens, we
tested the corresponding prediction that the observed distributions of
annual and perennial \emph{Heliophila} species would be significantly
associated with historic drought frequency. We tested for a relationship
between drought frequency and life history, season, and their
interaction by analysis of variance while including species as a random
effect using the lme4 package in R (Bates \emph{et al.}, 2014) and
compared annuals and perennials using Tukey adjusted post-hoc contrasts.
We next calculated the mean drought frequency during the winter, spring,
summer and autumn for each species. Because shared evolutionary history
of closely related species can lead to spurious associations between
traits and environments (Felsenstein, 1985), we tested for a
relationship between life history strategy and drought frequency while
controlling for phylogeny using phylogenetic logistic regression (Ives
\& Garland, 2010). This statistical approach is designed to control for
the confounding effects of common ancestry's influence on demographic
features such as geospatial relationships when addressing hypotheses
about the role of natural selection on trait distributions.

@@

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/phylogeny} \caption{Species and examples of herbaria specimens of
\emph{Heliophila} (a) Phylogeny and life history strategies of species
studied. Orange circles at branch tips mark annual species and blue
circles mark perennial species. At internal nodes, pie charts indicate
the estimated posterior probability of being annual versus perennial.
Example herbaria specimens accessed via GBIF of (b) \emph{H. minima},
(c) \emph{H. deserticola}, (d) \emph{H. coronopifolia} and (e) \emph{H.
ephemera}. Images (b,d,e) courtesy of The Bavarian Natural History
Collections (CC BY-SA 4.0) and (c) The London Natural History Museum (CC
BY 4.0). Links to images are found in the supplement.}\label{fig:phylogeny}
\end{figure}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/maps_boxplots} \caption{Locations of occurrence records of (a) annual and (b)
perennial \emph{Heliophila.} Drought frequency during the (c) winter,
(d) spring, (e) summer and (f) autumn detected using the VHI. (g)
Drought frequencies during each season at the observation locations of
annual and perennial \emph{Heliophila} (Post-hoc contrasts annuals and
perennials from ANOVA, ** = p \textless{} 0.01).}\label{fig:mapsboxplots}
\end{figure}

@@

After all filtering steps, 2192 records for 42 species (Figure
\ref{fig:phylogeny}, Table \ref{tab:speciesmeanstable}) passed for
further analyses. The number of samples varied between species, with a
mean of 52.19 samples per species. \emph{H. rigidiuscula} had the most
records, 201, and \emph{H. cornellsbergia} the fewest, 2 (Table
\ref{tab:speciesmeanstable}).

There were clear visual differences between the distributions of the 960
annual and the 1232 perennial \emph{Heliophila} observation records
(Figure \ref{fig:mapsboxplots}, Figure \ref{fig:speciesmaps}). While
annual species were generally found in the western regions of South
Africa and Namibia, primarily in the Cape Floristic Region and Succulent
Karoo (Figure \ref{fig:mapsboxplots}a), the occurrence of perennials
extended to the southern and eastern coast of South Africa (Figure
\ref{fig:mapsboxplots}b).

@@

We found that the frequency of drought was significantly higher at the
locations of occurrence records for annual species. When comparing
across all occurrence records (all records rather than species means,
Figure \ref{fig:mapsboxplots}g), a mixed-model analysis of variance
which included species as random effect revealed a significant
relationship between drought frequency and life history (p \textless{}
0.01), season (p \textless{} 0.01) and their interaction (p \textless{}
0.01) (Table \ref{tab:anovatable}). Post-hoc contrasts showed that the
frequency of drought was significantly higher at the location of annuals
during the summer (z ratio = 3.93, p = \textless{} 0.01), and autumn (z
ratio = 4.06, p \textless{} 0.01). Because a comparison across all
occurrence records does not account for variation in the number of
records per species (Table \ref{tab:speciesmeanstable}) or species
relatedness (Figure \ref{fig:phylogeny}a), we also tested whether mean
drought frequency values of each species were significantly different
between annuals and perennials using phylogenetic logistic regression.
We found that the mean drought frequencies were significantly higher
(\(\alpha = 0.05\)) in annual species during the spring, summer, and
autumn (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a,b).
These findings indicate that common ancestry alone does not explain
differences in the drought frequencies experienced between the
environments of annual and perennial \emph{Heliophila}.

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:modelstable}Phylogenetic logistic regressions between life history, and the mean drought frequency observed at specimen sites of Heliophila species the winter, spring, summer, and autumn}
\begin{tabular}{lll}
\toprule
Predictor & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{P}\\
\midrule
Intercept & 0.7231 & 0.6636\\
Winter drought freq. & -1.5452 & 0.7274\\ \midrule
Intercept & 5.0107 & 0.0534\\
Spring drought freq. & -12.9014 & 0.0464\\ \midrule
Intercept & 7.7093 & 0.0054\\
Summer drought freq. & -19.9056 & 0.0042\\ \midrule
Intercept & 7.0162 & 0.0082\\
Fall drought freq. & -20.8174 & 0.0067\\ \midrule
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Annual species were scored as 0 and perennial species as 1.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/line_and_dates} \caption{(a) Comparison (mean +- SE) of drought frequency across
seasons calculated at the occurrence locations of GBIF records of annual
and perennial species of \emph{Heliophila}. (b) Results from
phylogenetic logistic regression, where the line shows the model fit and
* = p \textless{} 0.05, ** = p \textless{} 0.01. Annuals were scored as
0 and perennials as 1. (c) Collection dates of GBIF records of annual
and perennial species of \emph{Heliophila} in relation to the
photoperiodic calendar where day 1 is intermediate to the autumn equinox
and winter solstice.}\label{fig:lineplots}
\end{figure}

The preceding results indicate that annual species are found in
environments where droughts are significantly more frequent, especially
in the summer and autumn. Classic life history theory hypothesizes that
annuality reflects adaptation to such environments because it allows
species to escape predictable stressful conditions. If this is the case,
we would expect that annuals spend the drought prone seasons of summer
and autumn as seeds. To test this hypothesis, we compared the dates of
occurrence records between annual and perennial \emph{Heliophila}
species. The distributions reveal a considerable difference in the
timing of observation of these two life histories. In comparison to
perennials, which appear to be collected throughout the year, annuals
are almost exclusively observed during the winter and spring (Figure
\ref{fig:lineplots}b). The differences between the distribution of
collection dates were significant by all tests (ks.test D = 0.25, p
\textless{} 0.01; bartlett.test K2 = 503.18, p \textless{} 0.01) This is
consistent with a model of life history in which annual species flower
in the spring, set seed, senesce, and die before the summer. Thus, these
annual species are likely to remain dormant during the summer and
autumn, when drought is the strongest predictor of the distributions of
annual and perennial life histories (Figure \ref{fig:lineplots}a).

@@

To test the hypothesis that annual and perennial plants reflect
adaptation to alternative drought environments we examined the
distributions of life history strategies in the large and diverse
mustard genus, \emph{Heliophila}. Using metadata of 2192 occurrence
records and a 34 year dataset of satellite-detected droughts, we tested
the prediction that annual species are more often observed in
drought-prone locations than perennial species, when controlling for
phylogenetic relatedness. We found that drought frequency is
significantly different between the distributions of annual and
perennial species, with annuals being found in environments with more
frequent drought, and that this signal is strongest during the seasons
when annuals are likely escaping via seed dormancy. These results remain
significant while controlling for the phylogenetic relationships of
\emph{Heliophila} species, yielding support for the role that natural
selection has played in driving contemporary distributions of these
alternatives strategies in relation to drought regimens.

@@

In conclusion, we find strong support for classic life history theory
that predicts that annuality is adaptive in environments with frequent
and predictable droughts. We report evidence consistent with a life
history model in annuals in which they escape drought prone seasons
during the seed phase of their life cycle. We also find evidence that
the distributions of perennial lineages may indicate a competitive
advantage in areas where droughts are infrequent. More broadly, this
work highlights the irreplaceable value of natural history collections
and demonstrates the power of combining such information with large
scale remote sensing data to address outstanding classic hypotheses in
ecology and evolution.

@@

\leavevmode\hypertarget{ref-bates2014fitting}{}%
\textbf{\textnormal{Bates D}, \textnormal{Mächler M}, \textnormal{Bolker
B}, \textnormal{Walker S}}. \textbf{2014}. Fitting linear mixed-effects
models using lme4. \emph{arXiv preprint arXiv:1406.5823}.

@@

\leavevmode\hypertarget{ref-clausen1948experimental}{}%
\textbf{\textnormal{Clausen J}, \textnormal{Keck DD}, \textnormal{Hiesey
WM}, \textnormal{others}}. \textbf{1948}. Experimental studies on the
nature of species. III. Environresponses of climatic races of achillea.
\emph{Experimental studies on the nature of species. III.
Environresponses of climatic races of Achillea.}

\leavevmode\hypertarget{ref-cody1975ecology}{}%
\textbf{\textnormal{Cody ML}, \textnormal{MacArthur RH},
\textnormal{Diamond JM}, \textnormal{others}}. \textbf{1975}.
\emph{Ecology and evolution of communities}. Harvard University Press.

@@

\leavevmode\hypertarget{ref-dittberner2018natural}{}%
\textbf{\textnormal{Dittberner H}, \textnormal{Korte A},
\textnormal{Mettler-Altmann T}, \textnormal{Weber AP},
\textnormal{Monroe G}, \textnormal{Meaux J de}}. \textbf{2018}. Natural
variation in stomata size contributes to the local adaptation of
water-use efficiency in arabidopsis thaliana. \emph{Molecular ecology}
\textbf{27}: 4052--4065.

@@

\leavevmode\hypertarget{ref-raunkiaer1905types}{}%
\textbf{\textnormal{Raunkiaer CC}}. \textbf{1905}. \emph{Types
biologiques pour la géographie botanique}.

@@

\leavevmode\hypertarget{ref-turesson1925plant}{}%
\textbf{\textnormal{Turesson G}}. \textbf{1925}. The plant species in
relation to habitat and climate: Contributions to the knowledge of
genecological units. \emph{Hereditas} \textbf{6}: 147--236.
